Debiased Calibration for Generalized Two-Phase Sampling

### 2024-04-30 final/reports/prelim_paper_ov.tex
Changed DC to G(w_i) instead of d_i G(w_i / d_i). It works better.

  $$\hat Y_{\ell}(\hat \lambda, \phi^*) = \hat Y_{DCE}(\hat \lambda) + 
  \left(\sum_{i \in A_1} w_{1i} z_i q_i - \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(\hat
  \lambda) z_i q_i\right)\phi^*$$

@@

$$
\begin{aligned}
X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
X_{2i} &\stackrel{ind}{\sim} Unif(0, 4) \\
X_{3i} &\stackrel{ind}{\sim} N(0, 1) \\
\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
\pi_{1i} &= \min(\Phi_3(-x_{3i} - 2), 0.7) \\
\pi_{2i|1} &= \min(\Phi_3(-x_{1i} + 1), 0.9).
\end{aligned}
$$

where $\Phi_3$ is the CDF of a t-distribution with 3 degrees of freedom.
%\textcolor{red}{The simulation setup is somewhat strange. The first order
%inclusion probability for the second-phase sampling, $\pi_{2i \mid 1}$, is a
%function of $x_{4i}$ which is completely independent of $(x_{1i}, x_{2i}, y_i)$.
%In this case, your second-phase sampling is essentially equivalent to the simple
%random sampling. }
This is a two-phase extension of the setup in \cite{kwon2024debiased}. We
consider a finite population of size $N = 10,000$ with both the Phase 1 and
Phase 2 sampling occuring under Poisson (Bernoulli) sampling. This yields a
Phase 1 sample
size of $E[n_1] \approx 1100$ and a Phase 2 sample size of
$E[n_2] \approx 300$. In the Phase 1 sample, we observe 
$(X_1, X_2)$ while in the Phase 2 sample we observe $(X_1, X_2, Y)$. This
simulation does not deal with model misspecification, and we compare the
proposed method for the parameter $\bar Y_N$ with four approaches:

\begin{itemize}
  \item[1.] $\pi^*$-estimator: $\hat Y_{\pi^*} = N^{-1} \sum_{i \in A_2}
    \frac{y_i}{\pi_{1i} \pi_{2i|1}},$
  \item[2.] Two Phase Regression estimator (TP-Reg): 
    $\hat Y_{reg} = \sum_{i \in A_1} \frac{\bf x_i' \hat \beta}{\pi_{1i}} + 
    \sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}}(y_i - \bf x_i' \hat \beta)$ 
    where $\hat \beta = 
    \left(\sum_{i \in A_2} \bf x_i \bf x_i'\right)^{-1} \sum_{i \in A_2} \bf x_i y_i$
    and $\bf x_i = (x_{1i}, x_{2i})^T$,
    % \textcolor{red}{Dr. Kim, should I modify the simulation so that the
    % regression estimator also includes $g(\pi_{2i|1}^{-1})$ as a covariate?}
  \item[3.] Debiased Calibration with Population Constraints (DC-Pop): This 
    solves

@@

\begin{table}[ht!]
  \centering
  \begin{tabular}{lcc}
    \toprule
    Estimator & Estimated Variance & Notes \\
    \midrule
    $\pi^*$ & $N^{-2} \sum_{i \in A_2} \left(\pi_{2i|1}^{-2} - \pi_{2i|1}^{-1}\right)
    y_i^2 $ & \\
    TP-Reg  & 
    {\scriptsize $N^{-2}\left(\sum\limits_{i \in A_1} 
        \left(\frac{1 - \pi_{1i}}{\pi_{1i}^{2}}\right) \eta_i^2 
        + \sum\limits_{i \in A_2} \frac{1}{\pi_{1i} \pi_{2i|1}}(d_{2i|1} - 1)
        (y_i - \bf x_i' \hat \beta)^2 \right)$}
    & 
      {\scriptsize$\eta_i = \bf x_i \hat \beta + \frac{\delta_{2i}}{\pi_{2i|1}}(y_i - 
      \bf x_i \hat \beta)$} \\
      %$\hat \beta = \left(\sum_{i \in A_2} \bf x_i \bf x_i'\right)^{-1} \sum_{i \in A_2} \bf x_i y_i$ \\
    DC-Pop  & {\scriptsize$\sum\limits_{i \in A_2} c_i (y_i - q_i z_i \hat \gamma)^2 
              (w_{1i} \hat w_{2i|1}(\hat \lambda) / N)^2$}
            & {\scriptsize$c_i = ((1 - \pi_{1i}) \pi_{2i|1} + (1 - \pi_{2i|1}))$} \\
      DC-Est  & 
      \scriptsize{$N^{-2}\left(\sum\limits_{i \in A_1} \left(\frac{1 -
        \pi_{1i}}{\pi_{1i}^2}\right) \eta_i^2 + \sum\limits_{i \in A_2}
          \left(\frac{1}{\pi_{1i}\pi_{2i|1}}\right) (d_{2i|1} - 1) (y_i - q_i
          z_i \hat \phi)^2\right)$} & 
      {\scriptsize $\eta_i = q_i z_i \hat \phi +
          \frac{\delta_{2i|1}}{\pi_{2i|1}}(y_i - q_i z_i \hat \phi)$} \\
    \bottomrule
  \end{tabular}
  \caption{This table gives the formulas for each variance estimator used in
  Simulation 1. For the derivation of the estimated variance of the DC-Pop
estimatator see Appendix A.}
  \label{tab:varforms}
\end{table}

\textcolor{red}{Dr. Kim, please see Appendix A for my derivation of the variance
DC-Pop.}

%    Pi_dc <- 
%      diag((1 - p2_df$pi1) * p2_df$pi2 + (1 - p2_df$pi2)) * 
%      (w1i * dc_w / nrow(pop_df))^2
%    err <- p2_df$Y - t(gam_dc) %*% t(z_dc) * qi
%    vhat <- (err) %*% Pi_dc %*% t(err)

@@

\begin{table}[ht!]
  \centering
\input{tables/tpdcsim_mean.tex}
% This table was generated from /src/explore/20240411-tpdcsim.qmd
\caption{This table shows the results of Simulation Study 1. It displays the
Bias, RMSE, empirical 95\% confidence interval, and a t-statistic assessing the
unbiasedness of each estimator for the estimators: $\pi^*$, TP-Reg, DC-Pop, and
DC-Est.}
\label{tab:tpdc-mean}
\end{table}

@@

\textcolor{red}{I am a little worried about the second term of $\hat{T}$. It is somewhat weird to assume that $T_g= \sum_{i \in U} g(d_{2i}) q_i$ is known. One simple way is to use a HT estimator  of $T_g$ from sample $A_2$. That is, use $\hat{T}_{g, {\rm HT}} = \sum_{i \in A_2} d_{2i} g( d_{2i}) q_i$ or even use a GREG estimator of $T_g$. 
%I wonder we can assume that an unbiased estimator of $T_g$ is available from $A_c=A_1 \cup A_2$. That is, let's assume that we have $\hat{T}_c = W \hat{T}_1 + (1- W) \hat{T}_2$ as the estimator  for $T$. 
}

>>>>

  $$\hat Y_{\ell}(\lambda, \phi^*) = \hat Y_{DCE}(\hat \lambda) + 
  \left(\sum_{i \in A_1} w_{1i} z_i q_i - \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(
  \lambda) z_i q_i\right)\phi^*$$

@@

$$
\begin{aligned}
X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
X_{2i} &\stackrel{ind}{\sim} Unif(0, 4) \\
X_{3i} &\stackrel{ind}{\sim} N(0, 1) \\
\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
%\pi_{1i} &= \min(\Phi_3(-x_{3i} - 2), 0.7) \\
\pi_{1i} &= n_1 / N \\
\pi_{2i|1} &= \min(\Phi_3(-x_{1i} + 1), 0.9).
\end{aligned}
$$

where $\Phi_3$ is the CDF of a t-distribution with 3 degrees of freedom.
%\textcolor{red}{The simulation setup is somewhat strange. The first order
%inclusion probability for the second-phase sampling, $\pi_{2i \mid 1}$, is a
%function of $x_{4i}$ which is completely independent of $(x_{1i}, x_{2i}, y_i)$.
%In this case, your second-phase sampling is essentially equivalent to the simple
%random sampling. }
This is a two-phase extension of the setup in \cite{kwon2024debiased}. We
consider a finite population of size $N = 10,000$ with the Phase 1 sampling
being a simple random sample (SRS) and the 
Phase 2 sampling occuring under Poisson sampling. This yields a
Phase 1 sample
size of $n_1 = 1000$ and a Phase 2 sample size of
$E[n_2] \approx 267$. In the Phase 1 sample, we observe 
$(X_1, X_2)$ while in the Phase 2 sample we observe $(X_1, X_2, Y)$. This
simulation does not deal with model misspecification, and we compare the
proposed method for the parameter $\bar Y_N$ with four approaches:

\begin{itemize}
  \item[1.] $\pi^*$-estimator: $\hat Y_{\pi^*} = N^{-1} \sum_{i \in A_2}
    \frac{y_i}{\pi_{1i} \pi_{2i|1}},$
  \item[2.] Two Phase Regression estimator (TP-Reg): 
    $\hat Y_{reg} = \sum_{i \in A_1} \frac{\bf x_i' \hat \beta}{\pi_{1i}} + 
    \sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}}(y_i - \bf x_i' \hat \beta)$ 
    where $\hat \beta = 
    \left(\sum_{i \in A_2} \bf x_i \bf x_i'\right)^{-1} \sum_{i \in A_2} \bf x_i y_i$
    and $\bf x_i = (x_{1i}, x_{2i})^T$,
    % \textcolor{red}{Dr. Kim, should I modify the simulation so that the
    % regression estimator also includes $g(\pi_{2i|1}^{-1})$ as a covariate?}

    \textcolor{red}{You should include an intercept in $\bf x_i$. }
  \item[3.] Debiased Calibration with Population Constraints (DC-Pop): This 
    solves

@@

\begin{table}[ht!]
  \centering
  \begin{tabular}{lcc}
    \toprule
    Estimator & Estimated Variance & Notes \\
    \midrule
    $\pi^*$ 
    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_Y + 
      \sum\limits_{i \in A_2} \frac{w_{1i}^2}{N^2} 
        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}y_i^2$}
    & {\scriptsize$\hat s^2_Y = \frac{1}{n_2 - 1}\sum\limits_{i \in A_2} (y_i -
    \bar y_i)^2$} \\
    TP-Reg  
    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_\eta + 
      \sum\limits_{i \in A_2} \frac{w_{1i}}{N^2} 
        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
    & {\scriptsize$ \eta_i = x_i'\beta + \delta_{2i} w_{2i|1}\varepsilon_i,\;
        \varepsilon_i = (y_i - x_i' \beta)$} \\
    DC-Pop  
    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_\varepsilon + 
      \sum\limits_{i \in A_2} \frac{w_{1i}}{N^2} 
        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
    & {\scriptsize$\eta_i = z_i' \phi + \delta_{2i}w_{2i|1}\varepsilon_i,\;
        \varepsilon_i = (y_i - z_i'\phi)$} \\
    DC-Est  
    & {\scriptsize$\left(\frac{1}{n_1} - \frac{1}{N}\right) \hat s^2_\eta + 
      \sum\limits_{i \in A_2} \frac{w_{1i}}{N^2} 
        \frac{(1 - \pi_{2i|1})}{\pi_{2i|1}^2}\varepsilon_i^2$}
    & {\scriptsize$\eta_i = z_i' \phi + \delta_{2i}w_{2i|1}\varepsilon_i,\;
        \varepsilon_i = (y_i - z_i' \phi)$} \\
    \bottomrule
  \end{tabular}
  \caption{This table gives the formulas for each variance estimator used in
  Simulation 1. For the derivation of the estimated variance of the DC-Pop
estimatator see Appendix A.}
  \label{tab:varforms}
\end{table}

%\textcolor{red}{I am not sure about the variance estimation formula in Table 1. They are different from what we discussed before. I wonder whether you can make the sampling rate $n/N$ negligible so that the second term becomes insignificant. How about making $N=50,000$, $n_1=500$, and $n_2=200$?}

@@

\begin{table}[ht!]
  \centering
\input{tables/tpdcsim_mean.tex}
% This table was generated from ../../src/explore/proto-dctpmcmc.R
\caption{This table shows the results of Simulation Study 1. It displays the
Bias, RMSE, empirical 95\% confidence interval, and a t-statistic assessing the
unbiasedness of each estimator for the estimators: $\pi^*$, TP-Reg, DC-Pop, and
DC-Est.}
\label{tab:tpdc-mean}
\end{table}

\textcolor{red}{Dr. Kim is this confidence interval good enough? It seems
slightly high for me.}

\textcolor{red}{I also give the mean variance estimates, the Monte Carlo mean,
as well as a test for bias.}
\begin{table}[ht!]
  \centering
{\color{red} \input{tables/tpdcsim_var.tex} }
% This table was generated from ../../src/explore/proto-dctpmcmc.R
\caption{\textcolor{red}{This table shows the variance estimates from Simulation
    Study 1. It displays the Monte Carlo variance of $\hat Y^{(b)}$, the average
    estimated variance, the variance of the variance estimator, and a
    t-statistic assessing the unbiasedness of each variance estimator.}}
\label{tab:tpdc-var}
\end{table}

@@

\textcolor{red}{I am a little worried about the second term of $\hat{T}$. It is
  somewhat weird to assume that $T_g= \sum_{i \in U} g(d_{2i}) q_i$ is known.
  One simple way is to use a HT estimator  of $T_g$ from sample $A_2$. That is,
  use $\hat{T}_{g, {\rm HT}} = \sum_{i \in A_2} d_{2i} g( d_{2i}) q_i$ or even
  use a GREG estimator of $T_g$. 
%I wonder we can assume that an unbiased estimator of $T_g$ is available from
  %$A_c=A_1 \cup A_2$. That is, let's assume that we have $\hat{T}_c = W
  %\hat{T}_1 + (1- W) \hat{T}_2$ as the estimator  for $T$. 
}
